---
title: "MoneyPuck"
author: "Dan Savickas"
date: "4/30/2017"
output:
  html_document:
    toc: yes
    toc_float: yes
subtitle: A Statistical Examination of Player Quality in the NHL
---

##Abstract

If any professional sports league draws little or no attention to contract details, it is the NHL. The league receives paltry coverage in the media and is generally regarded as the least popular of the major sports in the United States. Thus, poorly drawn up contracts, grossly overpaid and underpaid players, and dark horses generally fly under the radar to the casual fan. However, even a brief analysis of the data suggests that there is a significant discrepancy between the players that produce wins for their team, and those that are the most highly paid. Of course, there is the unquantifiable star power factor to consider, but for the purposes of winning games, NHL front offices appear to be off-base. In the spirit of the sabermetrics that revolutionized the MLB, and produced the Oscar-nominated picture, "MoneyBall", this inquiry is an effort to develop an algorithm that more accurately reflects player value, than any existing metric.

##Acknowledgements

Special thanks are owed to Professor Sean Carver, PhD, of American University, for providing guidance and his own knowledge of RStudio, which greatly contributed to this final product. Also providing invaluable assistance in this inquiry was Paul Jeffries, an American University student, and financial engineer at Fannie Mae. Statistics referenced were pulled from hockey-reference .com and salary data from sportrac. com.

***

```{r setup, include=FALSE, comment = NA}
knitr::opts_chunk$set(echo = FALSE)

# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))

library(readr)
library(corrplot)
NHL_Skaters <- read_csv("~/Documents/NHL_Stats.csv")
View(NHL_Skaters)
##Skaters

NHL_Goalies <- read_csv("~/Documents/NHL_Stats_Goalies.csv")
View(NHL_Goalies)
##Goalies

summary(NHL_Skaters)
summary(NHL_Goalies)
##Summary Statistics by position

'+/-mean' <- (mean(NHL_Skaters$`+/-`))
'+/-sd' <- (sd(NHL_Skaters$`+/-`))
##Coding mean and SD for single variables, in this case +/-

z.score <- function(x){
  (x-mean(x))/sd(x)
}
##Coding z-score as a function

z.score(NHL_Skaters$`+/-`)
##Using that function to create z-scores for a single variable

scale(NHL_Skaters$`+/-`)
##Different code to find the z-scores of a variable


NHL_Skaters$Salary <- c(0.6, 0.675, 0.8325, 0.8325, 0.8325, 0.8325, 0.775, 2.6, 0.8325, 1.7,
                           0.9, 3.25, 1, 0.8325, 0.7, 0.65, 0.8325, 0.66, 4, 0.575, 0.825,
                           3, 5, 1.75, 0.8325, 3.1, 4, 3, 3.8, 6, 7.875, 9.25, 3.2, 0.65,
                           0.8325, 0.874125, 0.95, 0.8075, 0.675, 0.725, 1, 2.25, 0.8325,
                           0.8325, 0.8325, 2.4, 3, 0.8325, 0.8325, 0.625, 2.5, 0.7, 2, 5.375,
                           0.8325, 6, 1, 0.8325, 0.675, 0.7, 0.8075, 0.8, 0.6, 0.6, 0.7, 2,
                           2, 0.8, 4, 3, 2.5, 0.75, 0.9, 0.6, 0.625, 0.9, 0.75, 0.9, 5, 5,
                           1.1, 4, 8.75, 7.25, 0.8325, 5, 0.65, 0.8325, 0.575, 0.575, 0.575,
                           0.8325, 0.6, 0.7, 0.8325, 0.8325, 0.775, 0.6, 0.65, 4.5, 3.9, 0.75,
                           0.6, 0.65, 0.6525, 5.25, 0.95, 2.65, 1.15, 2, 1.25, 2.25, 3, 2,
                           6, 5.4, 6, 0.8325, 7.5, 0.8325, 0.7, 0.6, 0.65, 0.65, 0.575, 0.635,
                           4.5, 0.6, 0.75, 0.95, 0.675, 0.75, 0.775, 1.1, 1.25, 0.8, 0.7,
                           0.8325, 0.6, 0.8325, 5.5, 2.5, 0.8325, 2.5, 6, 4, 0.8325, 6, 0.6,
                           0.8325, 0.8325, 0.65, 0.6325, 0.7, 2.825, 0.6, 0.6, 2.2, 5, 3.5,
                           1, 0.687333, 2.9, 5, 5.85, 5.4, 0.8325, 0.65, 5.5, 5.75, 0.925,
                           3.5, 0.625, 0.7, 0.8325, 0.575, 0.65, 0.8325, 0.6125, 0.575, 0.6,
                           0.65, 2.2, 2.75, 6, 3, 0.8, 4.5, 0.95, 0.8325, 4.8375, 0.95, 4.75,
                           5, 0.8325, 5.5, 3.575, 6.5, 3.75, 0.575, 0.8, 0.6, 0.75, 0.8325,
                           0.675, 0.8325, 0.6075, 0.575, 0.575, 0.75, 0.575, 2, 1.5, 3.6,
                           0.8325, 0.8325, 6.875, 0.875, 4.55, 4, 6, 10.5, 0.8325, 10.5,
                           0.874125, 0.575, 0.8, 0.8325, 0.675, 1.65, 0.8325, 0.575, 0.8,
                           0.8325, 1.9, 2.5, 0.8325, 0.8, 2, 4.75, 6, 4.5, 0.65, 2.4, 1.3,
                           5.5, 5, 0.8325, 6, 5, 0.6, 0.8325, 0.625, 0.8325, 0.9, 0.8325,
                           0.67, 0.675, 4, 0.91875, 0.65, 2, 3.65, 3.5, 0.625357, 0.6, 0.71,
                           5, 0.875, 2.2, 0.8325, 0.775, 2.5, 7.5, 5.75, 6, 0.8325, 0.6,
                           0.6, 0.65, 0.6, 0.575, 0.6, 0.575, 0.575, 0.6, 1.025, 4, 0.715,
                           4, 1.875, 5.5, 0.6, 0.675, 4, 5, 0.575, 0.8325, 6, 0.8325, 5, 2.75,
                           4.25, 7.75, 0.925, 0.585, 0.8325, 3.5, 1.05, 0.725, 0.95, 1.85,
                           0.8325, 0.8325, 0.8325, 0.725, 3.1, 4, 0.8325, 0.8325, 3, 1.5,
                           6.5, 1.8, 3, 2, 6, 6, 6, 0.8325, 0.8325, 0.705, 0.8325, 0.575,
                           0.8325, 0.65, 3, 0.6, 0.8325, 0.575, 1.2, 3.35, 1.3, 1.25, 0.8325,
                           0.8325, 0.95, 4, 3.5, 4, 3.5, 6.35, 4, 0.75, 4.5, 3.75, 1.6,
                           0.8325, 0.6, 2.5, 1.55, 1.4, 0.8325, 0.575, 0.925, 1, 0.575, 1.25,
                           0.6, 0.65, 6.075, 0.63, 2, 4, 3.9, 7, 3.9, 7.45, 0.735833, 10,
                           7, 0.8325, 0.64, 0.6, 0.65, 0.65, 0.8325, 0.925, 0.6, 0.65, 0.8325,
                           0.85, 0.6, 3.75, 0.65, 1.2, 4, 1.05, 2.375, 4.25, 9, 9, 6.25, 2,
                           2.6, 2.66667, 5.4, 3.5, 3.2, 0.8325, 0.55, 0.6, 0.6, 0.575, 0.8325,
                           0.8325, 0.575, 0.6, 0.575, 0.8325, 0.625, 4.2, 0.95, 1.3, 1, 6,
                           7, 0.8325, 4.5, 3, 4.25, 0.875, 7.857143, 1, 2.85, 5.75, 5, 0.635,
                           0.575, 0.625, 0.8325, 0.7, 0.675, 0.6, 0.6, 0.575, 0.75, 0.8, 0.7,
                           1, 2.5, 0.8325, 0.75, 0.625, 1.25, 1.1, 5, 0.8325, 0.725, 1.75,
                           0.8325, 0.6, 5, 4.5, 6.5, 6, 4.5, 0.575, 0.575, 0.575, 0.575, 0.65,
                           0.575, 0.575, 0.6, 0.575, 1.833333, 0.8425, 0.575, 0.6125, 0.575,
                           0.575, 0.575, 0.8325, 0.575, 3, 3.5, 4, 2, 4, 2.5, 9, 5, 4, 4.25,
                           6, 6, 0.575, 0.8325, 0.8, 0.575, 0.8325, 0.575, 0.575, 0.8325,
                           0.8325, 4.875, 0.575, 0.8, 2.2, 3.5, 1, 7, 4.25, 0.8325, 2.5, 2.5,
                           2, 5.5, 2, 2.5, 4.5, 3.5, 3.7, 6, 0.65, 0.6, 0.6, 1, 5, 0.6, 2.75,
                           5, 1.1, 0.7, 0.8325, 1, 0.8325, 1.65, 3.25, 8, 0.8325, 1.65, 4.7,
                           2.1, 4, 8, 2.5, 4.5, 0.925, 0.7, 0.75, 0.6, 0.65, 4.5, 0.575, 0.65,
                           0.575, 0.8325, 0.725, 0.874125, 1.1, 1.5, 0.7, 4.9, 0.9, 2.25,
                           0.8, 0.8, 7, 6.5, 0.75, 2.35, 1.1, 4, 3.75, 4, 3.8, 7, 0.925,
                           0.8325, 0.8325, 0.8325, 0.625, 0.95, 0.76815, 2.25, 0.8325, 0.725,
                           2.35, 0.85, 0.65, 0.725, 2.75, 0.84, 4.75, 4, 4, 3.35, 0.8325,
                           0.8325, 2, 0.8325, 4.3, 5.125, 9, 4.25, 0.7, 0.8325, 0.575, 0.575,
                           0.575, 0.575, 0.12776, 0.575, 0.6675, 0.575, 4.083, 0.575, 0.575,
                           0.8, 0.625, 2.7, 3.667, 2.3, 0.59, 0.62, 3.85, 1, 0.575, 7.25,
                           2.1, 4.75, 1.4, 0.575, 7, 9.5, 10.9, 0.675, 0.575, 0.7, 0.8325,
                           0.925, 0.6, 0.8325, 0.6, 2.75, 0.625, 3.8, 0.8325, 2.25, 0.575,
                           2.8, 1.85, 0.7, 5, 5.2, 4.25, 3, 6.5, 6.75, 6, 6, 5.76, 0.575,
                           0.6, 2.725, 0.575, 0.8325, 0.575, 1.05, 2.9, 2.5, 1, 0.65, 1.125,
                           0.7, 5.6, 0.9, 1, 0.9, 4.4, 0.8325, 3.7, 0.8325, 7, 3.75, 7,
                           6.5, 5.25, 8, 0.575, 1.25, 0.8325, 0.575, 0.575, 0.8325, 0.575,
                           0.8325, 0.575, 0.575, 6.5, 1.25, 0.6, 3.6, 0.8, 0.615, 3.7, 1.6,
                           8.5, 4.5, 1.9, 4.45, 0.575, 4, 4, 0.8325, 4.25, 4.25, 0.8325,
                           4, 0.8325, 1.2, 0.65, 0.65, 2.5, 0.575, 2.25, 1.2, 5, 0.925,
                           2.95, 0.8325, 0.65, 4.05, 4.2, 0.8325, 0.8325, 4.5, 5, 0.8325,
                           0.575, 0.6, 0.65, 0.8, 0.95, 0.65, 0.625, 0.8125, 0.296102,
                           0.8325, 0.650, 0.20556, 0.675, 1.25, 0.625, 0.27776, 0.2, 0.6,
                           1.25, 0.17776, 0.7, 0.14224, 1.1, 0.7825, 0.574986, 0.6, 3.75,
                           2.75, 0.9, 0.7425, 0.715, 0.6, 2, 0.6, 1.3, 2.6, 0.8, 0.735,
                           2.8, 4, 1.25, 1.55, 1, 2.1, 1.5, 0.725, 2.5, 1.1, 1, 2, 4.35,
                           0.335843, 2, 2.501625, 2.75, 1.25, 4.5, 3.5, 5, 1.3, 1, 4.25,
                           0.6325, 0.575, 0.8325, 0.575, 0.6, 0.575, 0.8325, 0.95, 0.6, 0.7,
                           2.5, 0.8325, 0.8325, 1.025, 3.5, 0.6, 0.6, 0.925, 0.7, 3.5, 3.1,
                           0.8325, 5, 0.8325, 6, 0.85, 3.3, 1.5, 7, 7, 0.8325, 0.8325, 0.575,
                           0.65, 0.6, 0.6, 0.874125, 2.7, 1.2, 5, 0.8325, 0.8325, 0.85,
                           2.125, 0.75, 0.925, 5.75, 0.8325, 0.875, 0.8325, 1, 2.5, 3, 5,
                           7, 0.8325, 0.8325, 5.8, 5, 0.8425, 0.575, 0.575, 0.575, 0.575,
                           0.575, 0.775, 0.8325, 2.8, 5.5, 0.875, 1.75, 0.85, 2, 4.25, 1.75,
                           2.75, 0.8325, 4, 5.75, 3.25, 4.5, 4.25, 3.4, 10, 7)
##Salary Data for all skaters

NHL_Goalies$Salary <- c(0.6, 1.5, 4.15, 6.5, 0.6, 0.6, 1, 0.8325, 7.5, 0.8325, 1.2, 0.8,
                           1, 3.15, 3.5, 0.7, 0.7, 2.55, 0, 0.7, 0.65, 8.5, 0.8325, 1.7,
                           0.8325, 2.7, 0.6, 6, 6, 0.675, 0.925, 0.65, 4.5, 6, 3.85, 5.5,
                           0.575, 4.166666, 0.8, 0.75, 5, 6.714, 1.5, 0.9, 7, 0.575, 0.65,
                           1.55, 5, 7, 0.95, 0.8325, 6, 0.75, 0.575, 7, 0.575, 4.75, 2, 0.675,
                           1, 9.5, 0.6, 1.35, 0.78, 4.75, 0.725, 1.75, 4.1, 0.575, 0.55,
                           5.75, 3, 0.575, 1.25, 0.70875, 2.2, 0.8325, 0.575, 0.75, 0.68,
                           5, 0.8, 0.575, 0.6, 5.95, 6, 1.6, 0.575, 4.75, 1, 0.6925, 0.575,
                           7, 0.85)
##Salary Data for all goalies

##Summary Stats for both sets of salary data

NHL_Skaters$TmPoints <- c(105, 105, 105, 105, 105, 105, 105, 105, 105, 105, 105, 105, 105,
                             105, 105, 105, 105, 105, 105, 105, 105, 105, 105, 105, 105,
                             105, 105, 105, 105, 105, 105, 105, 70, 70, 70, 70, 70, 70, 70,
                             70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70,
                             70, 70, 70, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95,
                             95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95,
                             95, 95, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78,
                             78, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78, 78,
                             78, 78, 78, 78, 78, 78, 87, 87, 87, 87, 87, 87, 87, 87, 87,
                             87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87,
                             87, 87, 87, 87, 108, 108, 108, 108, 108, 108, 108, 108, 108,
                             108, 108, 108, 108, 108, 108, 108, 108, 108, 108, 108, 108,
                             108, 108, 108, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94,
                             94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94,
                             109, 109, 109, 109, 109, 109, 109, 109, 109, 109, 109, 109,
                             109, 109, 109, 109, 109, 109, 109, 109, 109, 109, 109, 109, 
                             109, 48, 48, 48, 48, 48, 48, 48, 48, 48, 48, 48, 48, 48, 48,
                             48, 48, 48, 48, 48, 48, 48, 48, 48, 48, 48, 48, 79, 79, 79,
                             79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79,
                             79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79,
                             79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79, 79,
                             79, 79, 79, 79, 79, 79, 103, 103, 103, 103, 103, 103, 103,
                             103, 103, 103, 103, 103, 103, 103, 103, 103, 103, 103, 103,
                             103, 103, 103, 103, 103, 103, 103, 103, 81, 81, 81, 81, 81,
                             81, 81, 81, 81, 81, 81, 81, 81, 81, 81, 81, 81, 81, 81, 81,
                             81, 81, 81, 81, 81, 86, 86, 86, 86, 86, 86, 86, 86, 86, 86,
                             86, 86, 86, 86, 86, 86, 86, 86, 86, 86, 86, 86, 86, 86, 86,
                             106, 106, 106, 106, 106, 106, 106, 106, 106, 106, 106, 106,
                             106, 106, 106, 106, 106, 106, 106, 106, 106, 106, 106, 106,
                             106, 106, 106, 106, 103, 103, 103, 103, 103, 103, 103, 103,
                             103, 103, 103, 103, 103, 103, 103, 103, 103, 103, 103, 103,
                             103, 103, 103, 103, 103, 103, 103, 103, 70, 70, 70, 70, 70,
                             70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 70,
                             70, 70, 70, 70, 70, 70, 70, 70, 70, 70, 94, 94, 94, 94, 94,
                             94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94,
                             94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94,
                             94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94,
                             94, 94, 94, 94, 94, 94, 94, 94, 102, 102, 102, 102, 102, 102,
                             102, 102, 102, 102, 102, 102, 102, 102, 102, 102, 102, 102,
                             102, 102, 102, 102, 102, 102, 98, 98, 98, 98, 98, 98, 98, 98,
                             98, 98, 98, 98, 98, 98, 98, 98, 98, 98, 98, 98, 98, 98, 98,
                             98, 98, 98, 98, 98, 98, 98, 88, 88, 88, 88, 88, 88, 88, 88,
                             88, 88, 88, 88, 88, 88, 88, 88, 88, 88, 88, 88, 88, 88, 88,
                             88, 88, 88, 88, 88, 111, 111, 111, 111, 111, 111, 111, 111,
                             111, 111, 111, 111, 111, 111, 111, 111, 111, 111, 111, 111,
                             111, 111, 111, 111, 111, 111, 111, 111, 111, 111, 111, 99, 99,
                             99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99,
                             99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99,
                             99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99, 99,
                             99, 99, 99, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94,
                             94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 94, 95,
                             95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95, 95,
                             95, 95, 95, 95, 95, 94.3, 103, 102.6, 82.7, 93, 92.55, 99.9, 94,
                             103, 71, 80.28, 96.5, 91.2, 78, 82.6, 89, 97.6, 97.75, 72, 85,
                             85.3, 55, 84.3, 111.7, 95, 87, 85, 87, 100.5, 99, 71.2, 81.5,
                             110.9, 77, 72.5, 98.5, 73.5, 61, 103, 77, 80, 78, 80, 89, 89.7,
                             97, 92.5, 83.5, 83, 79, 90.5, 81, 94.3, 94, 57, 73, 74.5, 79.5,
                             90, 80.5, 85.5, 103.5, 69, 69, 69, 69, 69, 69, 69, 69, 69, 69,
                             69, 69, 69, 69, 69, 69, 69, 69, 69, 69, 69, 69, 69, 69, 69, 69,
                             69, 69, 69, 69, 69, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87,
                             87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87, 87,
                             87, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118,
                             118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118, 118,
                             118)
##Points (standings) for each player's team
##Players who played for more than one got the weighted average depending on how many games
##they played for each

na.omit(NHL_Goalies)
na.omit(NHL_Skaters)

# adding first MRs

Goalies <- NHL_Goalies$GS > 12
#Defining Goalies as those who started more than 12 games

NHL_Goalies$SVPG <- NHL_Goalies$SV/NHL_Goalies$GP
#Creation of the SavePerGame statistic

NHL_Goalies$WinPerc <- NHL_Goalies$W/NHL_Goalies$GS
#Creation of the Win Percentage statistic for goalies

NHL_Goalies$SOPerc <- NHL_Goalies$SO/NHL_Goalies$GS
#Creation of the Shutout Percentage statistic for goalies

# adding first complex MRs

# MASS PASTE
Age.z <- scale(NHL_Skaters$Age)
Games.z <- scale(NHL_Skaters$GP)
Goals.z <- scale(NHL_Skaters$G)
Assists.z <- scale(NHL_Skaters$A)
PlusMinus.z <- scale(NHL_Skaters$`+/-`)
PenMins.z <- scale(NHL_Skaters$PIM)
Shots.z <- scale(NHL_Skaters$S)
ShotPerc.z <- scale(NHL_Skaters$`S%`)
IceTime.z <- scale(NHL_Skaters$TOI)
TimePerGame.z <- scale(NHL_Skaters$ATOI)
Blocks.z <- scale(NHL_Skaters$BLK)
Hits.z <- scale(NHL_Skaters$HIT)
FaceOffs.z <- scale(NHL_Skaters$`FO%`)
FaceOffWins.z <- scale(NHL_Skaters$FOW)
SkaterSalary.z <- scale(NHL_Skaters$Salary)
TeamPoints.z <- scale(NHL_Skaters$TmPoints)
#Z-scores for all relevant skater data

GoalieAge.z <- scale(NHL_Goalies$Age)
Starts.z <- scale(NHL_Goalies$GS)
Wins.z <- scale(NHL_Goalies$W)
Saves.z <- scale(NHL_Goalies$SV)
SavePerc.z <- scale(NHL_Goalies$`SV%`)
GoalsAgainstAvg.z <- scale(NHL_Goalies$GAA)
Shutouts.z <- scale(NHL_Goalies$SO)
ShutoutPerc.z <- scale(NHL_Goalies$SOPerc)
GoalieSalary.z <- scale(NHL_Goalies$Salary)
SavePerGame.z <- scale(NHL_Goalies$SVPG)
QualityStartPerc.z <- scale(NHL_Goalies$`QS%`)
#Z-scores for all relevant goalie data
RightWing <- NHL_Skaters$GP > 16 & NHL_Skaters$Pos=='RW'
LeftWing <- NHL_Skaters$GP > 16 & NHL_Skaters$Pos=='LW'
Center <- NHL_Skaters$GP > 16 & NHL_Skaters$Pos=='C'
Forward <- RightWing | LeftWing | Center
#Defining Forward as an LW, C, or RW playing more than 16 games
Defender <- NHL_Skaters$GP > 16 & NHL_Skaters$Pos=='D'
#Defining Defender as a D playing more than 16 Games
Other <- !Forward & !Defender
# Define Others as skaters who played fewer than 16 games (neither forwards nor defenders in this set)


GoalieSalarySVPG.MR <- summary(lm(NHL_Goalies$Salary[Goalies] ~ NHL_Goalies$Age[Goalies] + NHL_Goalies$GS[Goalies]
                                    + NHL_Goalies$WinPerc[Goalies] + NHL_Goalies$SVPG[Goalies]
                                    + NHL_Goalies$SOPerc[Goalies] + NHL_Goalies$`QS%`[Goalies]))

GoalieWinsGAA.MR <- summary(lm(NHL_Goalies$WinPerc[Goalies] ~ NHL_Goalies$Age[Goalies]
                               + NHL_Goalies$GAA[Goalies]  
                               + NHL_Goalies$SOPerc[Goalies] + NHL_Goalies$`QS%`[Goalies]))

GoalieSalarySVPerc.MR <- summary(lm(NHL_Goalies$Salary[Goalies] ~ NHL_Goalies$Age[Goalies] + NHL_Goalies$GS[Goalies]
                              + NHL_Goalies$WinPerc[Goalies] + NHL_Goalies$`SV%`[Goalies]
                              + NHL_Goalies$SOPerc[Goalies] + NHL_Goalies$`QS%`[Goalies]))

GoalieWinsSVPG.MR <- summary(lm(NHL_Goalies$WinPerc[Goalies] ~ NHL_Goalies$Age[Goalies]
                                + NHL_Goalies$SVPG[Goalies]
                                + NHL_Goalies$SOPerc[Goalies] + NHL_Goalies$`QS%`[Goalies]))

GoalieWinsSVPerc.MR <- summary(lm(NHL_Goalies$WinPerc[Goalies] ~ NHL_Goalies$Age[Goalies] + 
     NHL_Goalies$`SV%`[Goalies] + NHL_Goalies$SOPerc[Goalies] + 
     NHL_Goalies$`QS%`[Goalies]))

GoalieSalaryGAA.MR <- summary(lm(NHL_Goalies$Salary[Goalies] ~ NHL_Goalies$Age[Goalies] + NHL_Goalies$GS[Goalies]
                                 + NHL_Goalies$WinPerc[Goalies] + NHL_Goalies$GAA[Goalies]
                                 + NHL_Goalies$SOPerc[Goalies] + NHL_Goalies$`QS%`[Goalies]))

#Defining multiple regression summaries for Goalie Regressions
#Note: GAA and 'SV%' are cross-correlated

ForwardsSalaryShots.MR <- summary(lm(NHL_Skaters$Salary[Forward] ~ NHL_Skaters$Age[Forward] + NHL_Skaters$GP[Forward] 
                                     + NHL_Skaters$S[Forward] + NHL_Skaters$A[Forward] + NHL_Skaters$PIM[Forward] 
                                     + NHL_Skaters$`S%`[Forward] + NHL_Skaters$TOI[Forward] 
                                     + NHL_Skaters$BLK[Forward] + NHL_Skaters$HIT[Forward] + NHL_Skaters$FOW[Forward] 
                                     + NHL_Skaters$`+/-`[Forward]))

ForwardsSalaryGoals.MR <- summary(lm(NHL_Skaters$Salary[Forward] ~ NHL_Skaters$Age[Forward] + NHL_Skaters$GP[Forward] 
                                     + NHL_Skaters$G[Forward] + NHL_Skaters$A[Forward] + NHL_Skaters$PIM[Forward] 
                                     + NHL_Skaters$`S%`[Forward] + NHL_Skaters$TOI[Forward] 
                                     + NHL_Skaters$BLK[Forward] + NHL_Skaters$HIT[Forward] + NHL_Skaters$FOW[Forward] 
                                     + NHL_Skaters$`+/-`[Forward]))

ForwardsTmPointsShots.MR <- summary(lm(NHL_Skaters$TmPoints[Forward] ~ NHL_Skaters$Age[Forward] + NHL_Skaters$GP[Forward] 
                                       + NHL_Skaters$A[Forward] + NHL_Skaters$PIM[Forward] 
                                       + NHL_Skaters$S[Forward] + NHL_Skaters$`S%`[Forward] + NHL_Skaters$TOI[Forward] 
                                       + NHL_Skaters$BLK[Forward] + NHL_Skaters$HIT[Forward] + NHL_Skaters$FOW[Forward] 
                                       + NHL_Skaters$`+/-`[Forward]))

ForwardsTmPointsGoals.MR <- summary(lm(NHL_Skaters$TmPoints[Forward] ~ NHL_Skaters$Age[Forward] + NHL_Skaters$GP[Forward] 
                                       + NHL_Skaters$G[Forward] + NHL_Skaters$A[Forward] + NHL_Skaters$PIM[Forward] 
                                       + NHL_Skaters$`S%`[Forward] + NHL_Skaters$TOI[Forward] 
                                       + NHL_Skaters$BLK[Forward] + NHL_Skaters$HIT[Forward] + NHL_Skaters$FOW[Forward] 
                                       + NHL_Skaters$`+/-`[Forward]))
#Defining multiple regression summaries for forward regressions
#Note: Goals and Shots are cross-correlated

DefenderSalaryGoals.MR <- summary(lm(NHL_Skaters$Salary[Defender] ~ NHL_Skaters$Age[Defender] + NHL_Skaters$GP[Defender] 
                                     + NHL_Skaters$G[Defender] + NHL_Skaters$A[Defender] + NHL_Skaters$PIM[Defender] 
                                     + NHL_Skaters$`S%`[Defender] + NHL_Skaters$TOI[Defender] 
                                     + NHL_Skaters$BLK[Defender] + NHL_Skaters$HIT[Defender] + NHL_Skaters$`+/-`[Defender]))

DefenderSalaryShots.MR <- summary(lm(NHL_Skaters$Salary[Defender] ~ NHL_Skaters$Age[Defender] + NHL_Skaters$GP[Defender] 
                                     + NHL_Skaters$A[Defender] + NHL_Skaters$PIM[Defender] 
                                     + NHL_Skaters$S[Defender] + NHL_Skaters$`S%`[Defender] + NHL_Skaters$TOI[Defender] 
                                     + NHL_Skaters$BLK[Defender] + NHL_Skaters$HIT[Defender] + NHL_Skaters$`+/-`[Defender]))

DefenderTmPointsGoals.MR <- summary(lm(NHL_Skaters$TmPoints[Defender] ~ NHL_Skaters$Age[Defender] + NHL_Skaters$GP[Defender] 
                                       + NHL_Skaters$G[Defender] + NHL_Skaters$A[Defender] + NHL_Skaters$PIM[Defender] 
                                       + NHL_Skaters$`S%`[Defender] + NHL_Skaters$TOI[Defender] 
                                       + NHL_Skaters$BLK[Defender] + NHL_Skaters$HIT[Defender] 
                                       + NHL_Skaters$`+/-`[Defender]))

DefenderTmPointsShots.MR <- summary(lm(NHL_Skaters$TmPoints[Defender] ~ NHL_Skaters$Age[Defender] + NHL_Skaters$GP[Defender] 
                                       + NHL_Skaters$A[Defender] + NHL_Skaters$PIM[Defender] 
                                       + NHL_Skaters$S[Defender] + NHL_Skaters$`S%`[Defender] + NHL_Skaters$TOI[Defender] 
                                       + NHL_Skaters$BLK[Defender] + NHL_Skaters$HIT[Defender] 
                                       + NHL_Skaters$`+/-`[Defender]))
#Defining multiple regressions for defender regressions
#Note: Goals and Shots are cross-correlated

JeffrickasFWD <- (2.892*Age.z[Forward] + 8.703*Games.z[Forward] + 3.733*Goals.z[Forward]
                    + 6.605*Assists.z[Forward] - 0.559*PenMins.z[Forward] 
                    + 1.295*ShotPerc.z[Forward] + 18.36*Shots.z[Forward]
                    + 4.117*Blocks.z[Forward] + 0.902*Hits.z[Forward]
                    + 0.302*FaceOffWins.z[Forward] + 185.559*PlusMinus.z[Forward] - 11.1726)
#Beginning of aggregate value statistic, the 'Jeffrickas Index', for forwards
                      
JeffrickasDEF <- (0.3064*Age.z[Defender] + 2.621*Games.z[Defender] - 0.904*Goals.z[Defender] 
                    + 2.604*Assists.z[Defender] - 0.244*PenMins.z[Defender] + 3.287*ShotPerc.z[Defender] 
                    + 0.974*Shots.z[Defender] + 0.0463*Blocks.z[Defender] - 0.0163*Hits.z[Defender] 
                    + 122.191*PlusMinus.z[Defender] - 14.7354)
#Beginning of aggregate value statistic, the 'Jeffrickas Index', for defenders

JeffrickasGOL <- (4.38*SavePerc.z[Goalies] - 0.727*GoalieAge.z[Goalies] - 17.573*GoalsAgainstAvg.z[Goalies] 
                  + 0.569*QualityStartPerc.z[Goalies] + 2.97*ShutoutPerc.z[Goalies] - 2.538*SavePerGame.z[Goalies] - 2.733958)
#Beginning of aggregate value statistic, the 'Jeffrickas Index', for goalies

#Correlation Plots
string.free.goalies <- NHL_Goalies[,-c(1:2,4,13:14,18,20:22,28:29)]
string.free.skaters <- NHL_Skaters[,-c(1:4,5,27,20)]
Goalies.Cor <- cor(string.free.goalies)
Skaters.Cor <- cor(string.free.skaters)
corrplot(Goalies.Cor, method = "circle") #plot matrix
corrplot(Skaters.Cor, method = "circle") #plot matrix

#Correlations
cor(NHL_Skaters$Salary[Forward], NHL_Skaters$TmPoints[Forward])
cor(NHL_Skaters$Salary[Forward], JeffrickasFWD)
cor(NHL_Skaters$TmPoints[Forward], JeffrickasFWD)
cor(NHL_Skaters$`+/-`[Forward], NHL_Skaters$TmPoints[Forward])

cor(NHL_Skaters$Salary[Defender], NHL_Skaters$TmPoints[Defender])
cor(NHL_Skaters$Salary[Defender], JeffrickasDEF)
cor(NHL_Skaters$TmPoints[Defender], JeffrickasDEF)
cor(NHL_Skaters$`+/-`[Defender], NHL_Skaters$TmPoints[Defender])

cor(NHL_Goalies$Salary[Goalies], NHL_Goalies$WinPerc[Goalies])
cor(NHL_Goalies$Salary[Goalies], JeffrickasGOL)
cor(NHL_Goalies$WinPerc[Goalies], JeffrickasGOL)
cor(NHL_Goalies$GAA[Goalies], NHL_Goalies$WinPerc[Goalies])


```

##Statistical Testing

Given that, in the NHL, different teams have various schemes, gameplans, line changes, and strategies in play, the most direct possible comparisons of player value come at the Goaltender position. A goalie is either going to allow a shot to go in, or he isn't. Of course, being an NHL goaltender is much harder work than that simplification, but that is the essence of a goaltender's  duties. The relevant statistics considered to be expected to contribute to a goalie's salary were Age, Win Percentage, Save Percentage, Goals Against per Game, Saves Per Game, Shutout Percentage, and Quality Start Percentage (A quality start is a game in which a goalie has a save percentage higher than the league average, or one in which he saves 88.5% of shots, while also allowing fewer than 2 goals). Knowing logically that Save Percentage, Saves Per Game, and Goals Against Average would be cross-correlated, three separate multiple regressions were run.

***

<br>

## Goalie Salary Regression Tests {.tabset .tabset-fade .tabset-pills}

As is evident by the data, the only metrics that ended up correlating with a Goalie's salary were Age and Games Started. No other hockey statistic registered as significant. Win%, GAA, SV%, SVPG, and QS% all produced fairly large p-values. Results are shown below.

***

### Save Percentage

```{r, comment = NA, null_prefix=TRUE}

GoalieSalarySVPerc.MR

```

***

<br>

### Goals Against Average

```{r, comment = NA, null_prefix=TRUE}

GoalieSalaryGAA.MR

```

*** 

<br>

### Saves Per Game

```{r, comment = NA, null_prefix=TRUE}

GoalieSalarySVPG.MR

```


***

<br>

## Goalie Win% Regression Tests {.tabset .tabset-fade .tabset-pills}

To test whether this is a reasonable, or effective, way to pay Goalies, a round of multiple regressions was run with Win Percentage as the dependent variable. This is to see if any hockey statistic can be statistically shown to have a significant correlation with wins. It will also show whether or not the significance levels were comparable to those for the salary regressions. Each statistic registered different significance levels. GAA had a p-value of about .0001, meaning there was just about a 99.99% chance that that statistic is significant for wins. SV% and SO% (in one of three tests) also registered statistically significant results. Full results are shown below.

***

### Save Percentage

```{r, comment = NA, null_prefix=TRUE}

GoalieWinsSVPerc.MR

```

***

<br>

### Goals Against Average

```{r, comment = NA, null_prefix=TRUE}

GoalieWinsGAA.MR

```

*** 

<br>

### Saves Per Game

```{r, comment = NA, null_prefix=TRUE}

GoalieWinsSVPG.MR

```

***

<br>

## The Jeffrickas Index

Some of the above results were striking, as GAA, SV%, and SO% registered significantly for Win%, but did not register as significant for salary data. This was a common finding, indicating that contribution to wins on the ice did not necessarily translate into a higher salary from the front office. The results from the Win% regressions led to the creation of an algorithm that could serve as an index to model how efficiently a Goalie contributes to his team winning. This model was generated by multiplying the z-score for each statistic (no. of standard deviations above the league average) by the square of the average t-value for that same statistic. That number was then adjusted to produce an average score of zero.

This index was named the "Jeffrickas Index" to combine my last name, with that of Paul Jeffries, whose contributions helped establish this inquiry. Here are the summary statistics of that index for Goalies:

```{r, comment = NA, null_prefix=TRUE}

summary(JeffrickasGOL)

```

<br>

When plotted, you can see that there is a definite positive correlation between a Goalie's Jeffrickas score, and win percentage. The Jeffrickas Index for Goalies correlated 73.33% with win percentage, making it more predictive than any one statistic (the next closest was GAA, which correlated slightly above 71%). 

```{r, comment = NA, null_prefix=TRUE}

plot(JeffrickasGOL, NHL_Goalies$WinPerc[Goalies])

```

***

<br>

## Forward Salary Regression Tests {.tabset .tabset-fade .tabset-pills}

Having established an index for Goalies, and a general understanding of what statistics contribute to both salary and Win%, the same methodology was then applied to Skaters (non-goalies). However, due to the differing roles each position plays on the ice, the Skaters were divided into Forwards and Defensemen, each getting their own regression model and Jeffrickas Index.

For Skaters, Shots (S) and Goals (G) were found to be cross-correlated, as it stands to reason that a player who takes more shots will score more goals. As such, two separate regressions were run to account for said cross correlation. For Forwards, much more correlation was found between relevant statistics and salary. Only assists (A), +/-, and Penalty Minutes (PIM) were never significantly correlated.

***

<br>

### Shots

```{r, comment = NA, null_prefix=TRUE}

ForwardsSalaryShots.MR

```

***

<br>

### Goals

```{r, comment = NA, null_prefix=TRUE}

ForwardsSalaryGoals.MR

```


***

<br>

## Forward Team Pts Regression Test {.tabset .tabset-fade .tabset-pills}

Similar regressions were run, but this time the dependent variable was Team Points (in the standings), as Skaters do not have win-loss records like Goalies. In these tests, +/- showed up as massively significant, with a negligible p-value in both tests. Assists came up as significant, as did Shots (S), Shot Percentage (S%), Time on Ice (TOI), and Blocked Shots (BLK).

***

<br>

### Shots

```{r, comment = NA, null_prefix=TRUE}

ForwardsTmPointsShots.MR

```

***

<br>

### Goals

```{r, comment = NA, null_prefix=TRUE}

ForwardsTmPointsGoals.MR

```


***

<br>

## Correlation Plots {.tabset .tabset-fade .tabset-pills}

### Skaters
```{r, comment = NA, null_prefix=TRUE}

corrplot(Skaters.Cor, method = "circle") #plot matrix

```

### Goalies
```{r, comment = NA, null_prefix=TRUE}

corrplot(Goalies.Cor, method = "circle") #plot matrix
```



